---
title: "Boston Biomass Results"
author: "Drew Trlica"
date: "July 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "E:/FragEVI/")
library(data.table)
library(ggplot2)
library(raster)

```

```{r data_setup1, echo=FALSE}
## this is all ~7k live stem records from Morreale for local FIA sites
live <- read.csv("processed/fia.live.stem.dbh.growth.csv") 
live <- as.data.table(live)
live <- live[order(DIAM_T0, decreasing=F),]

## these are the growth~dbh relationships (ca. 900) for all the Andy forest records, including psuedoreplicates calculated in 5 year intervals for each of the 195 tree cores taken
andy.bai <- read.csv("processed/andy.bai.dbh.pseudo.csv") 
andy.bai <- as.data.table(andy.bai)
andy.bai <- andy.bai[order(dbh.start, decreasing = F),]

## these are the 2006/2014 revisit records for the 6k street trees in Boston
street <- read.csv("processed/street.dbh.growth.results.csv") 
street <- as.data.table(street)
street <- street[order(dbh.2006, decreasing = F),]
```

## Estimates of annual biomass growth rates
Trees from all three sampling regimes showed a negative inverse relationship between stem DBH (cm) and relative annual growth rate (kg/kg) (Table 1). Median relative growth rates were higher in the urban forest and street tree samples than in FIA forest samples, and all stems displayed considerable intrasite variability, particularly in smaller stems. In the FIA rural forest samples, median DBH was `r round(live[,median(DIAM_T0, na.rm=T)], digits=1)` cm corresponding to median annual growth of `r round(live[, median(growth.ann.rel, na.rm=T)], digits=3)` kg/kg. Urban forest stems grew faster than rural forest, and showed higher growth in edge (<10m) compared to interior stems, with median DBH of `r round(andy.bai[seg.Edge=="E", median(avg.dbh, na.rm=T)], digits=1)` cm and `r round(andy.bai[seg.Edge=="I", median(avg.dbh, na.rm=T)], digits=1)` cm, respectively, corresponding to median annual growth of `r round(andy.bai[seg.Edge=="E", median(growth.mean, na.rm=T)], digits=3)` kg/kg and `r round(andy.bai[seg.Edge=="I", median(growth.mean, na.rm=T)], digits=3)` kg/kg. Street trees showed the highest median annual growth rate of `r round(street[record.good==1 & dbh.2006>5, median(npp.ann.rel, na.rm=T)], digits=3)` kg/kg corresponding to median DBH of `r round(street[record.good==1 & dbh.2006>5, median(dbh.2006, na.rm=T)], digits=1)`. The range and median stem DBH in each sample context were similar, but the street tree sample included very few conifers and a relatively large fraction of non-native taxa, primarily deciduous hardwoods including members of *Gleditsia*, *Zelkova* and *Pyrus*. 

```{r growth_dbh_mod, echo=FALSE}
### individual stem growth~dbh
## FIA dbh records
mod.fia.stem.nls <- summary(nls(growth.ann.rel ~ exp(a + b * log(DIAM_T0)), data=live, start=list(a=0, b=0)))
a.fia <- mod.fia.stem.nls$coefficients[1]
b.fia <- mod.fia.stem.nls$coefficients[2]
live$pred <- exp(a.fia+(b.fia*log(live$DIAM_T0)))

### Andy dbh records ## can get away with log-log model here because no negative growth records
# mod.int.edge.fin <- summary(lm(log(growth.mean)~log(avg.dbh)+seg.Edge, data=andy.bai)) # not including pseudoreps
mod.int.edge.fin <- summary(lm(log(biom.rel.ann)~log(dbh.start)+seg.Edge, data=andy.bai))
b0.bai <- mod.int.edge.fin$coefficients[1]
b1.bai <- mod.int.edge.fin$coefficients[2]
b2.bai <- mod.int.edge.fin$coefficients[3]
andy.bai[seg.Edge=="E", pred:=exp(b0.bai+(b1.bai*log(dbh.start)))]
andy.bai[seg.Edge=="I", pred:=exp(b0.bai+b2.bai+(b1.bai*log(dbh.start)))]

## street tree dbh records
mod.street.nls <- summary(nls(npp.ann.rel ~ exp(a + b * log(dbh.2006)), data=street[record.good==1 & dbh.2006>5,], start=list(a=0, b=0))) 
a.street <- mod.street.nls$coefficients[1]
b.street <- mod.street.nls$coefficients[2]
street$pred <- exp(a.street+(b.street*log(street$dbh.2006)))
```


```{r dbh_vs_stemgrowth_FIG, echo=FALSE, message=FALSE, fig.width=8.5, fig.height=3.7}
par(mfrow=c(1,3), mar=c(4,4,3,1))
main.xlim <- c(4, 80)
main.ylim <- c(-0.15, 1)
## FIA stem growth~dbh
col.type <- c("green", "forestgreen")
plot(live$DIAM_T0, live$growth.ann.rel, col=col.type[as.numeric(live$type)],
     pch=15, cex=0.3, xlim=main.xlim, ylim=main.ylim,
     xlab="Stem DBH (cm)", ylab="Growth rate (kg/kg)", main="FIA")
lines(live$DIAM_T0, live$pred,
       col="black", lwd=3)
legend(x=30, y=0.6, bty="n", legend = c("Hardwood", "Softwood"), fill=c("green", "forestgreen"))
abline(v=live[, median(DIAM_T0)])
abline(h=live[, median(growth.ann.rel)])

### Andy forest growth~dbh
col.edge <- c("royalblue", "purple")
# plot(andy.bai$avg.dbh, andy.bai$growth.mean, col=col.edge[as.numeric(as.factor(andy.bai$seg.Edge))],
#      pch=15, cex=0.3, xlim=main.xlim, ylim=main.ylim,
#      ylab="Growth Rate (kg/kg)", xlab="Stem DBH (cm)", main="Urban Forest")
# points(andy.bai$avg.dbh, exp(b0.bai)*exp(b1.bai*log(andy.bai$avg.dbh)), 
#        col=col.edge[1], pch=13, cex=0.4)
# points(andy.bai$avg.dbh, exp(b0.bai+b2.bai)*exp(b1.bai*log(andy.bai$avg.dbh)), 
#        col=col.edge[2], pch=13, cex=0.4)
### with pseudoreps
plot(andy.bai$dbh.start, andy.bai$biom.rel.ann, col=col.edge[as.numeric(as.factor(andy.bai$seg.Edge))],
     pch=15, cex=0.3, xlim=main.xlim, ylim=main.ylim,
     ylab="Growth Rate (kg/kg)", xlab="Stem DBH (cm)", main="Urban Forest")
lines(andy.bai[seg.Edge=="E", dbh.start], andy.bai[seg.Edge=="E", pred],
       col=col.edge[1], lwd=3)
lines(andy.bai[seg.Edge=="I", dbh.start], andy.bai[seg.Edge=="I", pred],
       col=col.edge[2], lwd=3)
legend(x=30, y=0.6, legend=c("Edge (<10m)", "Interior"), fill=col.edge, bty="n")
abline(v=andy.bai[seg.Edge=="E", median(dbh.start, na.rm=T)], col=col.edge[1])
abline(v=andy.bai[seg.Edge=="I", median(dbh.start, na.rm=T)], col=col.edge[2])
abline(h=andy.bai[seg.Edge=="E", median(biom.rel.ann, na.rm=T)], col=col.edge[1])
abline(h=andy.bai[seg.Edge=="I", median(biom.rel.ann, na.rm=T)], col=col.edge[2])

## street tree growth~dbh
plot(street[record.good==1 & dbh.2006>5, dbh.2006], street[record.good==1, npp.ann.rel],
     pch=15, col="red", cex=0.3, xlim=main.xlim, ylim=main.ylim,
     xlab="Stem DBH (cm)", ylab="Growth rate (kg/kg)", main="Street trees")
lines(street[record.good==1 & dbh.2006>5, dbh.2006], 
       street[record.good==1 & dbh.2006>5, pred],
       col="black", lwd=3)
abline(v=street[record.good==1 & dbh.2006>5, median(dbh.2006)])
abline(h=street[record.good==1 & dbh.2006>5, median(npp.ann.rel)])
```

##### **Figure 1**: Stem DBH and relative growth rate for different forest contexts. Note: Street tree figure truncated to exclude relatively small number of large DBH and large relative growth points.


Talk about the model construction of stem growth as function of dbh here
```{r Table_1, echo=FALSE}
require(knitr)
fia.mod <- mod.fia.stem.nls
andy.mod <- mod.int.edge.fin
street.mod <- mod.street.nls

context <- c("FIA (Hard/Softwood)", "Urban Forest (Edge/Int)", "Street Trees")

### number of stem observations
N <- c(length(live[PlotID%in%(live.plot[num.stems>20, PlotID]), DIAM_T0]),
       length(andy.bai[dbh.start>5, dbh.start]),
       length(street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", dbh.2006]))

### DBH ranges
dbh.med <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(DIAM_T0, probs=c(0.5))],
             andy.bai[is.finite(dbh.start), quantile(dbh.start, probs=c(0.5))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(dbh.2006, probs=c(0.5))]), digits=1)

dbh.low <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(DIAM_T0, probs=c(0.025))],
             andy.bai[is.finite(dbh.start),quantile(dbh.start, probs=c(0.025))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(dbh.2006, probs=c(0.025))]), digits=1)  

dbh.hi <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(DIAM_T0, probs=c(0.975))],
             andy.bai[is.finite(dbh.start),quantile(dbh.start, probs=c(0.975))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(dbh.2006, probs=c(0.975))]), digits=1)

dbh.col <- paste(dbh.med, " (", dbh.low, "-", dbh.hi, ")", sep="")

## growth ranges
growth.med <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(growth.ann.rel, probs=c(0.5))],
             andy.bai[is.finite(dbh.start),quantile(biom.rel.ann, probs=c(0.5))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(npp.ann.rel, probs=c(0.5))]), digits=3)

growth.low <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(growth.ann.rel, probs=c(0.025))],
             andy.bai[is.finite(dbh.start),quantile(biom.rel.ann, probs=c(0.025))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(npp.ann.rel, probs=c(0.025))]), digits=3)  

growth.hi <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(growth.ann.rel, probs=c(0.975))],
             andy.bai[is.finite(dbh.start),quantile(biom.rel.ann, probs=c(0.975))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(npp.ann.rel, probs=c(0.975))]), digits=3)

growth.col <- paste(growth.med, " (", growth.low, "-", growth.hi, ")", sep="")

### stem regression coefficients: growth = exp(a+b*log(dbh))
b0.mods <- round(c(fia.mod$coefficients[1], andy.mod$coefficients[1], street.mod$coefficients[1]), digits=2)
b1.mods <- round(c(fia.mod$coefficients[2], andy.mod$coefficients[2], street.mod$coefficients[2]), digits=2)
b2.mods <- c("NS", round(andy.mod$coefficients[3], digits=2), "NA")
b0.se <- round(c(fia.mod$coefficients[1,2], andy.mod$coefficients[1,2], street.mod$coefficients[1,2]), digits=2)
b1.se <- round(c(fia.mod$coefficients[2,2], andy.mod$coefficients[2,2], street.mod$coefficients[2,2]), digits=2)
b2.se <- c("", round(andy.mod$coefficients[3,2], digits=2), "")
b0.col <- paste(b0.mods, " (", b0.se, ")", sep="")
b1.col <- paste(b1.mods, " (", b1.se, ")", sep="")
b2.col <- paste(b2.mods, " (", b2.se, ")", sep="")

stem.tab <- data.frame(cbind(context, N, dbh.col, growth.col, b0.col, b1.col, b2.col))

kable(stem.tab, format="markdown", row.names=F, 
      col.names=c("Site", "N", "DBH (cm)", "Growth (kg/kg)", "B0", "B1", "B2"),
      align="l", caption=c("Stem growth rates in local site samples"))
```

```{r areal_growth, echo=FALSE, message=FALSE, warning=FALSE}
### FIA areal basis
live <- read.csv("processed/fia.live.stem.dbh.growth.csv")
live <- as.data.table(live)
live.plot <- live[,
                  .(length(unique(SubID)),
                    sum(biom.delt.spp, na.rm=T), 
                    sum(biom0.spp, na.rm=T),
                    length(DIAM_T0)),
                  by=PlotID]
names(live.plot) <- c("PlotID", "num.subplots", "biom.growth", "total.biom0.kg", "num.stems")
subplot.area <- (7.3152^2)*pi ## subplots are 24ft in radius
live.plot[,total.biom0.Mg.ha:=((total.biom0.kg/1000)/(num.subplots*subplot.area))*1E4] ## biomass density in aggregated area covered by the fully forested subplots
live.plot[,biom.growth.ann:=biom.growth/4.8]
live.plot[,biom.growth.ann.rel:=(biom.growth.ann)/total.biom0.kg]
hwood <- live[type=="H", .(sum(biom1.spp-biom0.spp, na.rm=T), 
                           sum(biom0.spp, na.rm=T)), 
              by=PlotID]
names(hwood)[2:3] <- c("biom.growth.spp.hw", "total.biom0.spp.kg.hw")
swood <- live[type=="S", .(sum(biom1.spp-biom0.spp, na.rm=T), 
                           sum(biom0.spp, na.rm=T)),
              by=PlotID]
names(swood)[2:3] <- c("biom.growth.spp.sw", "total.biom0.spp.kg.sw")

live.plot <- merge(live.plot, hwood, by="PlotID")
live.plot <- merge(live.plot, swood, by="PlotID")
live.plot[,biom.growth.ann.hw:=(biom.growth.spp.hw/4.8)/total.biom0.spp.kg.hw]
live.plot[,biom.growth.ann.sw:=(biom.growth.spp.sw/4.8)/total.biom0.spp.kg.sw]


## Andy plot-level
andy.dbh <- read.csv("processed/andy.dbh.proc.results.csv")
andy.dbh <- as.data.table(andy.dbh)
g <- andy.dbh[, .(sum(growth.kg), sum(growth.kg.ps), sum(biom)), by=.(seg, Plot.ID)] ## total biomass gain and total biomass for each plot
names(g)[3:5] <- c("growth.kg", "growth.kg.ps", "biom")
g[,rel.gain:=growth.kg/biom] ## this deals in forest growth as a function of biomass, not of forest area
g[,rel.gain.ps:=growth.kg.ps/biom]
g[seg==10, seg.F:="E"]
g[seg!=10, seg.F:="I"]
g[,biom.kg.ha:=(biom/(10*20))*1E4] ## get things in a standard biomass density
g[,mean(rel.gain), by=seg.F] ## 4.34% edge vs. 2.85% interior, annual gain on biomass per edge category
g[,mean(rel.gain.ps), by=seg.F] ## 3.91% edge vs. 2.74% interior, annual gain on biomass per edge category

```

Growth rates on an areal basis (kg-biomass/Mg-biomass/ha) were best predicted as an exponential function of log-biomass density (kg-biomass/ha), comparable to the growth-DBH relationships shown on a per-stem basis (Table 2). On an areal basis, the FIA rural forest plots showed a lower median hardwood growth rate of `r round(live.plot[num.stems>20, median(growth.ann.rel.hw, na.rm=T)], digits=3)` kg/Mg/ha corresponding to median biomass density of `r round(live.plot[num.stems>20, (median(total.biom0.spp.kg, na.rm=T)/675)*1E-3*1E4], digits=1)` Mg/ha, reflecting the lower per-stem rate of growth in the rural forest sample. In the FIA rural forest sample growth rates in individual stems did not differ significantly between hard- and softwoods, but given the comparatively low frequency of softwoods noted in the urban forest and street tree samples the the modeled areal-basis growth rate of hardwoods alone was used as the basis for estimating productivity in the city of Boston.
Areal growth rates in urban forest plots were estimated by estimating edge- and interior-specific stem growth rates of every tree >5cm DBH in each plot based on stem DBH. Median areal growth rate in the urban forest plots, accounting for edge position, was `r round(g[seg.F=="E", median(rel.gain, na.rm=T)], digits=3)` kg/Mg/ha in the edge (<10m) plots compared to `r round(g[seg.F=="I", median(rel.gain, na.rm=T)], digits=3)` kg/Mg/ha in interior plots across a range of biomass density from `r round(g[, (min(total.biom.kg)/(20*30))*1E4*1E-3], digits=1)`--`r round(g[, (max(total.biom.kg)/(20*30))*1E4*1E-3], digits=1)` Mg/ha. Per-area growth in edge canopy of urban forests was higher than interior canopy, but declined more quickly with greater biomass density. Street tree growth on a per-area basis could not be estimated directly with the sample scheme employed, and was instead areal biomass productivity was predicted on the basis of estimated productivity in simulated aggregates of individual stems (Methods).

```{r Table_2, echo=F, message=F}
## areal biomass density
live.N <- dim(live.plot[num.stems<20,]) ## update once you figure how to groom this for partial
live.dens <- round(live.plot[num.stems>20, median(((total.biom0.spp.kg/675)*1E4*1E-3), na.rm=T)], digits=1)
live.dens <- paste(live.dens, " (", 
                   round(live.plot[num.stems>20, quantile(((total.biom0.spp.kg/675)*1E4*1E-3), probs=0.025)], digits=1),
                   "-",
                   round(live.plot[num.stems>20, quantile(((total.biom0.spp.kg/675)*1E4*1E-3), probs=0.975)], digits=1),
                   ")", sep="")

andy.dens <- round(c(g[seg.F=="E", median(biom.kg.ha)/1000],
                     g[seg.F=="I", median(biom.kg.ha)]/1000), digits=1)
andy.dens <- paste(andy.dens[1], " (",
                   round(g[seg.F=="E", quantile(biom.kg.ha, probs=0.025)/1000], digits=1),
                   "-",
                   round(g[seg.F=="E", quantile(biom.kg.ha, probs=0.975)/1000], digits=1),
                   ") / ",
                   andy.dens[2], " (",
                   round(g[seg.F=="I", quantile(biom.kg.ha, probs=0.025)/1000], digits=1),
                   "-",
                   round(g[seg.F=="I", quantile(biom.kg.ha, probs=0.975)/1000], digits=1),
                   ")", sep="")

biom.dens.col <- c(live.dens, andy.dens, "NA")

## areal biomass growth rate
biom.growth <- round(live.plot[num.stems>20, median(growth.ann.rel.hw, na.rm=T)], digits=3)
andy.growth <- round(c(g[seg.F=="E", median(rel.gain)],
                     g[seg.F=="I", median(rel.gain)]), digits=3)
biom.growth <- paste(biom.growth, " (", 
                   round(live.plot[num.stems>20, quantile(growth.ann.rel.hw, probs=0.025)], digits=3),
                   "-",
                   round(live.plot[num.stems>20, quantile(growth.ann.rel.hw, probs=0.975)], digits=3),
                   ")", sep="")
andy.growth <- paste(andy.growth[1], " (",
                   round(g[seg.F=="E", quantile(rel.gain, probs=0.025)], digits=3),
                   "-",
                   round(g[seg.F=="E", quantile(rel.gain, probs=0.975)], digits=3),
                   ") / ",
                   andy.growth[2], " (",
                   round(g[seg.F=="I", quantile(rel.gain, probs=0.025)], digits=3),
                   "-",
                   round(g[seg.F=="I", quantile(rel.gain, probs=0.975)], digits=3),
                   ")", sep="")

biom.growth.col <- c(biom.growth, andy.growth, "NA")

stem.tab <- cbind(stem.tab, biom.dens.col, biom.growth.col)
names(stem.tab) <- c("N", "DBH (cm)", "Stem growth (kg/kg)", 
                     "B0", "B1", "B2", "Plot biom. (Mg/ha)", 
                     "Plot growth, Mg/Mg")
stem.tab
```
##### **Table 1**: Summary of stem growth in FIA plots, urban forest, and street trees. Figures in parentheses represent median and central 95% of data; significant model coefficients (p<0.05) are shown with standard errors and apply to the formula growth = exp(B0 + B1*log(DBH) + B2), with B2 corresponding to hardwoods vs. softwoods in FIA plots and in edge vs. interior plots in the urban forest. 



### Estimates of NPP in Boston
```{r npp_estimates, echo=FALSE}
npp.dat <- as.data.table(read.csv("processed/npp.estimates.V1.csv"))
npp.tab <- npp.dat[aoi>800 & !is.na(lulc), .(round(sum(fia.empir.npp.kg.hw.ground, na.rm=T)/(1000*2), digits=0),
                              round(sum(fia.empir.npp.kg.hw.forest, na.rm=T)/(1000*2), digits=0),
                              round(sum(fia.empir.npp.kg.hw.perv, na.rm=T)/(1000*2), digits=0),
                              round(sum(hyb.npp, na.rm=T)/(1000*2), digits=0)), by=lulc]
npp.tab$lulc.name <- c("Dev", "Other Veg", "Forest", "Water", "HD Resid", "LD Resid")
names(npp.tab) <- c("lulc", "FIA (Ground)", "FIA (canopy)", "FIA (pervious)", "Street+Urban Forest", "Type")
npp.tab <- cbind(npp.tab[,2:5, with=F], npp.tab$Type)
npp.tot <- apply(npp.tab[,1:4, with=F], 2, sum)
npp.tot <- matrix(nrow=1, ncol=5, c(npp.tot, "Total"))
npp.tab <- as.matrix(npp.tab)
npp.tab <- rbind(npp.tab, npp.tot)
npp.tab <- as.data.frame(cbind(npp.tab[,5], npp.tab[,1:4]))
names(npp.tab)[1] <- "Type"
npp.tab
```
##### **Table 2**: Estimated annual NPP (Mg-biomass/yr) for different Land-use/Land-cover classes in city of Boston.

Blah blah look how much NPP there was in these places wow Figure 2

```{r figure2, echo=FALSE}
npp.dat <- as.data.table(read.csv("processed/npp.estimates.V1.csv"))
npp.dat[,biom.bin:=cut(bos.biom30m, seq(0, 52000, by = 500), right=FALSE, ordered_result=T)]
par(mfrow=c(1,2), mar=c(4,4,2,0))

##hybrid results, NPP by biomass density bin
contain.npp <- npp.dat[!is.na(bos.biom30m) & !is.na(lulc), sum(hyb.npp, na.rm=T), by=.(biom.bin, lulc)]
contain.npp <- contain.npp[order(biom.bin), ]
contain.npp[, bin.num:=as.numeric(biom.bin)]
names(contain.npp)[3] <- "tot.npp.kg"
col.tmp <- c("forestgreen", "gray55", "salmon", "gold", "chartreuse3")
plot(contain.npp[lulc==1, bin.num], contain.npp[lulc==1, tot.npp.kg/1000], 
     col=col.tmp[1], pch=15, type="l", lwd=3,
     ylim=c(0, 500), xlab="Biomass density (Mg/ha)", xaxt="n", ylab="Total NPP (Mg-biomass)", main="Hybrid estimates") 
axis(side = 1, at = seq(0,100, by=20), labels = c(0, 10000, 20000, 30000, 40000, 50000))
for(e in 2:5){
  lines(contain.npp[lulc==e, bin.num], contain.npp[lulc==e, tot.npp.kg/1000], col=col.tmp[e], pch=15, type="l", lwd=3)
}
legend(x = 60, y = 400, legend = c("Forest", "Developed", "HD Resid.", "LD Resid.", "Other veg."), fill=col.tmp, bty="n")

## FIA empirical, canopy basis
contain.npp <- npp.dat[!is.na(bos.biom30m) & !is.na(lulc), sum(fia.empir.npp.kg.hw.forest, na.rm=T), by=.(biom.bin, lulc)]
contain.npp <- contain.npp[order(biom.bin), ]
contain.npp[, bin.num:=as.numeric(biom.bin)]
names(contain.npp)[3] <- "tot.npp.kg"
plot(contain.npp[lulc==1, bin.num], contain.npp[lulc==1, tot.npp.kg/1000], 
     col=col.tmp[1], pch=15, type="l", lwd=3,
     ylim=c(0, 500), xlab="Biomass density (Mg/ha)", xaxt="n", ylab="Total NPP (Mg-biomass)", main="FIA (canopy) estimates") 
axis(side = 1, at = seq(0,100, by=20), labels = c(0, 10000, 20000, 30000, 40000, 50000))
for(e in 2:5){
  lines(contain.npp[lulc==e, bin.num], contain.npp[lulc==e, tot.npp.kg/1000], col=col.tmp[e], pch=15, type="l", lwd=3)
}
legend(x = 60, y = 400, legend = c("Forest", "Developed", "HD Resid.", "LD Resid.", "Other veg."), fill=col.tmp, bty="n")

```

