---
title: "Boston Biomass Results"
author: "Drew Trlica"
date: "July 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "E:/FragEVI/")
library(data.table)
library(ggplot2)
library(raster)

```

```{r data_setup1, echo=FALSE}
live <- read.csv("processed/fia.live.stem.dbh.growth.csv")
live <- as.data.table(live)
andy.bai <- read.csv("processed/andy.bai.dbh.results.csv")
andy.bai <- as.data.table(andy.bai)
street <- read.csv("processed/street.dbh.growth.results.csv")
street <- as.data.table(street)
```

```{r dbh_records, echo=FALSE}
### individual stem growth~dbh
## FIA dbh records
mod.fia.stem.nls <- nls(growth.ann.rel ~ exp(a + b * log(DIAM_T0)), data=live, start=list(a=0, b=0))
## this is most parsimonious model; log-log including Hard/softwood factor not significant
# summary(mod.fia.stem.nls)

### Andy dbh records
mod.int.edge.fin <- summary(lm(log(growth.mean)~log(avg.dbh)+seg.Edge, data=andy.bai)) ## can get away with log-log model here because no negative growth records
b0.bai <- mod.int.edge.fin$coefficients[1]
b1.bai <- mod.int.edge.fin$coefficients[2]
b2.bai <- mod.int.edge.fin$coefficients[3]

## street tree dbh records
mod.street.nls <- nls(npp.ann.rel ~ exp(a + b * log(dbh.2006)), data=street[record.good==1,], start=list(a=0, b=0)) 
grub <- summary(mod.street.nls)

```
## Estimates of annual biomass growth rates
Trees from all three sampling regimes showed a negative inverse relationship between stem DBH (cm) and relative annual growth rate (kg/kg) (Table). Measured relative growth rates were higher in the urban forest and street tree samples than in FIA forest samples, and all samples displayed considerable variability in growth rate, particularly in smaller stems. In the FIA rural forest samples, median DBH was `r round(live[,median(DIAM_T0, na.rm=T)], digits=1)` cm corresponding to growth of `r round(live[, median(growth.ann.rel, na.rm=T)], digits=3)` kg/kg. Urban forest stems grew faster than rural forest, and showed significantly higher growth in edge (<10m) compared to interior stems, with median DBH of `r round(andy.bai[seg.Edge=="E", median(avg.dbh, na.rm=T)], digits=1)` cm and `r round(andy.bai[seg.Edge=="I", median(avg.dbh, na.rm=T)], digits=1)` cm, respectively, corresponding to median relative growth of `r round(andy.bai[seg.Edge=="E", median(growth.mean, na.rm=T)], digits=3)` kg/kg and `r round(andy.bai[seg.Edge=="I", median(growth.mean, na.rm=T)], digits=3)` kg/kg. Street trees showed the largest median relative growth rate of `r round(street[record.good==1, median(npp.ann.rel, na.rm=T)], digits=3)` kg/kg corresponding to median DBH of `r round(street[record.good==1, median(dbh.2006, na.rm=T)], digits=1)`. The range and median stem DBH in each sample context were similar, but the street tree sample included very few conifers and a relatively large fraction of non-native taxa, primarily deciduous hardwoods including members of *Gleditsia*, *Zelkova* and *Pyrus*. 

```{r dbh_vs_stemgrowth_FIG, echo=FALSE, message=FALSE, fig.width=8.5, fig.height=3.7}
par(mfrow=c(1,3), mar=c(4,4,3,1))
main.xlim <- c(4, 80)
main.ylim <- c(-0.15, 1)
## FIA stem growth~dbh
col.type <- c("green", "forestgreen")
plot(live$DIAM_T0, live$growth.ann.rel, col=col.type[as.numeric(live$type)],
     pch=15, cex=0.3, xlim=main.xlim, ylim=main.ylim,
     xlab="Stem DBH (cm)", ylab="Growth rate (kg/kg)", main="FIA")
points(live$DIAM_T0, predict(mod.fia.stem.nls),col="black", pch=13, cex=0.4)
legend(x=30, y=0.6, bty="n", legend = c("Hardwood", "Softwood"), fill=c("green", "forestgreen"))
abline(v=live[, median(DIAM_T0)])
abline(h=live[, median(growth.ann.rel)])

### Andy forest growth~dbh
col.edge <- c("royalblue", "purple")
plot(andy.bai$avg.dbh, andy.bai$growth.mean, col=col.edge[as.numeric(as.factor(andy.bai$seg.Edge))],
     pch=15, cex=0.3, xlim=main.xlim, ylim=main.ylim,
     ylab="Growth Rate (kg/kg)", xlab="Stem DBH (cm)", main="Urban Forest")
points(andy.bai$avg.dbh, exp(b0.bai)*exp(b1.bai*log(andy.bai$avg.dbh)), 
       col=col.edge[1], pch=13, cex=0.4)
points(andy.bai$avg.dbh, exp(b0.bai+b2.bai)*exp(b1.bai*log(andy.bai$avg.dbh)), 
       col=col.edge[2], pch=13, cex=0.4)
legend(x=30, y=0.6, legend=c("Edge (<10m)", "Interior"), fill=col.edge, bty="n")
abline(v=andy.bai[seg.Edge=="E", median(avg.dbh)], col=col.edge[1])
abline(v=andy.bai[seg.Edge=="I", median(avg.dbh)], col=col.edge[2])
abline(h=andy.bai[seg.Edge=="E", median(growth.mean)], col=col.edge[1])
abline(h=andy.bai[seg.Edge=="I", median(growth.mean)], col=col.edge[2])

## street tree growth~dbh
plot(street[record.good==1, dbh.2006], street[record.good==1, npp.ann.rel],
     pch=15, col="red", cex=0.3, xlim=main.xlim, ylim=main.ylim,
     xlab="Stem DBH (cm)", ylab="Growth rate (kg/kg)", main="Street trees")
points(street[record.good==1, dbh.2006], 
       street[record.good==1, exp(grub$coefficients[1]+(grub$coefficients[2]*log(dbh.2006)))],
       pch=8, col="black", cex=0.4)
abline(v=street[record.good==1, median(dbh.2006)])
abline(h=street[record.good==1, median(npp.ann.rel)])
```

##### **Figure 1**: Stem DBH and relative growth rate for different forest contexts. Note: Street tree figure truncated to exclude relatively small number of large DBH and large relative growth points.


```{r areal_growth, echo=FALSE, message=FALSE, warning=FALSE}
### FIA areal basis
live.plot <- live[,.(sum(biom1.spp-biom0.spp, na.rm=T),
                     sum(biom0.spp, na.rm=T),
                     length(DIAM_T0)), by=PlotID] 
names(live.plot)[2:4] <- c("biom.growth.spp", "total.biom0.spp.kg", 
                           "num.stems")
hwood <- live[type=="H", .(sum(biom1.spp-biom0.spp, na.rm=T), 
                           sum(biom0.spp, na.rm=T)), 
              by=PlotID]
names(hwood)[2:3] <- c("biom.growth.spp.hw", "total.biom0.spp.kg.hw")
swood <- live[type=="S", .(sum(biom1.spp-biom0.spp, na.rm=T), 
                           sum(biom0.spp, na.rm=T)),
              by=PlotID]
names(swood)[2:3] <- c("biom.growth.spp.sw", "total.biom0.spp.kg.sw")
live.plot[,growth.ann.rel.spp:=(biom.growth.spp/4.8)/total.biom0.spp.kg]
mod.spp <- lm(log(growth.ann.rel.spp)~log(((total.biom0.spp.kg/675)*1E4)), data=live.plot)
roop <- (mod.spp) ## slightly sloppier, R2 0.19, coefficients about the same
live.plot <- merge(live.plot, hwood, by="PlotID")
live.plot <- merge(live.plot, swood, by="PlotID")
live.plot[,growth.ann.rel.hw:=(biom.growth.spp.hw/4.8)/total.biom0.spp.kg.hw]
live.plot[,growth.ann.rel.sw:=(biom.growth.spp.sw/4.8)/total.biom0.spp.kg.sw]

## Andy plot-level
andy.dbh <- read.csv("processed/andy.dbh.proc.results.csv")
andy.dbh <- as.data.table(andy.dbh)
g <- andy.dbh[, .(sum(growth.kg), sum(biom)), by=.(seg, Plot.ID)] 
g[,rel.gain:=V1/V2] 
g[seg==10, seg.F:="E"]
g[seg!=10, seg.F:="I"]
names(g)[3:4] <- c("plot.growth.kg", "total.biom.kg")
g[, total.biom.Mg.ha:=(total.biom.kg/(20*30))*1E4*1E-3]
# g[,mean(rel.gain), by=seg.F] ## 4.34% edge vs. 2.85% interior, annual gain on biomass per edge category

```

```{r Table_1, echo=FALSE}
fia.mod <- summary(mod.fia.stem.nls)
andy.mod <- mod.int.edge.fin
street.mod <- grub
context <- c("FIA (Hard/Softwood)", "Urban Forest (Edge/Int)", "Street Trees")

### number of stem observations
N <- c(length(live[PlotID%in%(live.plot[num.stems>20, PlotID]), DIAM_T0]),
       length(andy.bai[,avg.dbh]),
       length(street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", dbh.2006]))

### DBH ranges
dbh.med <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(DIAM_T0, probs=c(0.5))],
             andy.bai[,quantile(avg.dbh, probs=c(0.5))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(dbh.2006, probs=c(0.5))]), digits=1)

dbh.low <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(DIAM_T0, probs=c(0.025))],
             andy.bai[,quantile(avg.dbh, probs=c(0.025))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(dbh.2006, probs=c(0.025))]), digits=1)  

dbh.hi <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(DIAM_T0, probs=c(0.975))],
             andy.bai[,quantile(avg.dbh, probs=c(0.975))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(dbh.2006, probs=c(0.975))]), digits=1)

dbh.col <- paste(dbh.med, " (", dbh.low, "-", dbh.hi, ")", sep="")

## growth ranges
growth.med <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(growth.ann.rel, probs=c(0.5))],
             andy.bai[,quantile(growth.mean, probs=c(0.5))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(npp.ann.rel, probs=c(0.5))]), digits=3)

growth.low <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(growth.ann.rel, probs=c(0.025))],
             andy.bai[,quantile(growth.mean, probs=c(0.025))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(npp.ann.rel, probs=c(0.025))]), digits=3)  

growth.hi <- round(c(live[PlotID%in%(live.plot[num.stems>20, PlotID]), quantile(growth.ann.rel, probs=c(0.975))],
             andy.bai[,quantile(growth.mean, probs=c(0.975))],
             street[dbh.2006>5 & standing.dead.2014==0 & Health!="Poor", quantile(npp.ann.rel, probs=c(0.975))]), digits=3)

growth.col <- paste(growth.med, " (", growth.low, "-", growth.hi, ")", sep="")

### stem regression coefficients: growth = exp(a+b*log(dbh))
b0.mods <- round(c(fia.mod$coefficients[1], andy.mod$coefficients[1], street.mod$coefficients[1]), digits=2)
b1.mods <- round(c(fia.mod$coefficients[2], andy.mod$coefficients[2], street.mod$coefficients[2]), digits=2)
b2.mods <- c("NS", round(andy.mod$coefficients[3], digits=2), "NA")
b0.se <- round(c(fia.mod$coefficients[1,2], andy.mod$coefficients[1,2], street.mod$coefficients[1,2]), digits=2)
b1.se <- round(c(fia.mod$coefficients[2,2], andy.mod$coefficients[2,2], street.mod$coefficients[2,2]), digits=2)
b2.se <- c("", round(andy.mod$coefficients[3,2], digits=2), "")
b0.col <- paste(b0.mods, " (", b0.se, ")", sep="")
b1.col <- paste(b1.mods, " (", b1.se, ")", sep="")
b2.col <- paste(b2.mods, " (", b2.se, ")", sep="")

stem.tab <- data.frame(cbind(N, dbh.col, growth.col, b0.col, b1.col, b2.col))

## areal biomass density
biom.dens <- round(live.plot[num.stems>20, median(((total.biom0.spp.kg/675)*1E4*1E-3), na.rm=T)], digits=1)
andy.dens <- round(c(g[seg.F=="E", median(total.biom.Mg.ha)],
                     g[seg.F=="I", median(total.biom.Mg.ha)]), digits=1)
biom.dens <- paste(biom.dens, " (", 
                   round(live.plot[num.stems>20, quantile(((total.biom0.spp.kg/675)*1E4*1E-3), probs=0.025)], digits=1),
                   "-",
                   round(live.plot[num.stems>20, quantile(((total.biom0.spp.kg/675)*1E4*1E-3), probs=0.975)], digits=1),
                   ")", sep="")
andy.dens <- paste(andy.dens[1], " (",
                   round(g[seg.F=="E", quantile(total.biom.Mg.ha, probs=0.025)], digits=1),
                   "-",
                   round(g[seg.F=="E", quantile(total.biom.Mg.ha, probs=0.975)], digits=1),
                   ") / ",
                   andy.dens[2], " (",
                   round(g[seg.F=="I", quantile(total.biom.Mg.ha, probs=0.025)], digits=1),
                   "-",
                   round(g[seg.F=="I", quantile(total.biom.Mg.ha, probs=0.975)], digits=1),
                   ")", sep="")

biom.dens.col <- c(biom.dens, andy.dens, "NA")

## areal biomass growth rate
biom.growth <- round(live.plot[num.stems>20, median(growth.ann.rel.hw, na.rm=T)], digits=3)
andy.growth <- round(c(g[seg.F=="E", median(rel.gain)],
                     g[seg.F=="I", median(rel.gain)]), digits=3)
biom.growth <- paste(biom.growth, " (", 
                   round(live.plot[num.stems>20, quantile(growth.ann.rel.hw, probs=0.025)], digits=3),
                   "-",
                   round(live.plot[num.stems>20, quantile(growth.ann.rel.hw, probs=0.975)], digits=3),
                   ")", sep="")
andy.growth <- paste(andy.growth[1], " (",
                   round(g[seg.F=="E", quantile(rel.gain, probs=0.025)], digits=3),
                   "-",
                   round(g[seg.F=="E", quantile(rel.gain, probs=0.975)], digits=3),
                   ") / ",
                   andy.growth[2], " (",
                   round(g[seg.F=="I", quantile(rel.gain, probs=0.025)], digits=3),
                   "-",
                   round(g[seg.F=="I", quantile(rel.gain, probs=0.975)], digits=3),
                   ")", sep="")

biom.growth.col <- c(biom.growth, andy.growth, "NA")

stem.tab <- cbind(stem.tab, biom.dens.col, biom.growth.col)
names(stem.tab) <- c("N", "DBH (cm)", "Stem growth (kg/kg)", 
                     "B0", "B1", "B2", "Plot biom. (Mg/ha)", 
                     "Plot growth, Mg/Mg")
stem.tab
```
##### **Table 1**: Summary of stem growth in FIA plots, urban forest, and street trees. Figures in parentheses represent median and central 95% of data; significant model coefficients (p<0.05) are shown with standard errors and apply to the formula growth = exp(B0 + B1*log(DBH) + B2), with B2 corresponding to hardwoods vs. softwoods in FIA plots and in edge vs. interior plots in the urban forest. 

Biomass growth on an areal basis was lowest in the FIA plots with a median growth rate in hardwoods of `r round(live.plot[num.stems>20, median(growth.ann.rel.hw, na.rm=T)], digits=3)` kg/Mg/ha corresponding to median biomass density of `r round(live.plot[num.stems>20, (median(total.biom0.spp.kg, na.rm=T)/675)*1E-3*1E4], digits=1)` Mg/ha, reflecting to the lower per-stem rate of growth in the rural forest sample. Areal growth rates in urban forest plots were estimated by estimating edge- and interior-specific stem growth rates of every tree >5cm DBH in each plot based on stem DBH. Median areal growth rate in the urban forest plots was `r round(g[seg.F=="E", median(rel.gain, na.rm=T)], digits=3)` kg/Mg/ha in the edge (<10m) plots compared to `r round(g[seg.F=="I", median(rel.gain, na.rm=T)], digits=3)` kg/Mg/ha in interior plots across a range of biomass density from `r round(g[, (min(total.biom.kg)/(20*30))*1E4*1E-3], digits=1)`--`r round(g[, (max(total.biom.kg)/(20*30))*1E4*1E-3], digits=1)` Mg/ha. Urban forest plots showed a significant exponential relationship to log-biomass density, but variability at all levels of density remained high. Due to the lower number of individual plots and somewhat restricted range of biomass density sampled in the urban forest context, a single growth factor was used for edge and interior biomass in order to estimate NPP in urban forest and high-biomass pixels. Street trees were not sampled on an areal basis, necessitating the estimation of areal biomass productivity on the basis of estimated productivity in simulated aggregates of individual stems. 

### Estimates of NPP in Boston
```{r npp_estimates, echo=FALSE}
npp.dat <- as.data.table(read.csv("processed/npp.estimates.V1.csv"))
npp.tab <- npp.dat[aoi>800 & !is.na(lulc), .(round(sum(fia.empir.npp.kg.hw.ground, na.rm=T)/(1000*2), digits=0),
                              round(sum(fia.empir.npp.kg.hw.forest, na.rm=T)/(1000*2), digits=0),
                              round(sum(fia.empir.npp.kg.hw.perv, na.rm=T)/(1000*2), digits=0),
                              round(sum(hyb.npp, na.rm=T)/(1000*2), digits=0)), by=lulc]
npp.tab$lulc.name <- c("Dev", "Other Veg", "Forest", "Water", "HD Resid", "LD Resid")
names(npp.tab) <- c("lulc", "FIA (Ground)", "FIA (canopy)", "FIA (pervious)", "Street+Urban Forest", "Type")
npp.tab <- cbind(npp.tab[,2:5, with=F], npp.tab$Type)
npp.tot <- apply(npp.tab[,1:4, with=F], 2, sum)
npp.tot <- matrix(nrow=1, ncol=5, c(npp.tot, "Total"))
npp.tab <- as.matrix(npp.tab)
npp.tab <- rbind(npp.tab, npp.tot)
npp.tab <- as.data.frame(cbind(npp.tab[,5], npp.tab[,1:4]))
names(npp.tab)[1] <- "Type"
npp.tab
```
##### **Table 2**: Estimated annual NPP (Mg-biomass/yr) for different Land-use/Land-cover classes in city of Boston.

Blah blah look how much NPP there was in these places wow Figure 2

```{r figure2, echo=FALSE}
npp.dat <- as.data.table(read.csv("processed/npp.estimates.V1.csv"))
npp.dat[,biom.bin:=cut(bos.biom30m, seq(0, 52000, by = 500), right=FALSE, ordered_result=T)]
par(mfrow=c(1,2), mar=c(4,4,2,0))

##hybrid results, NPP by biomass density bin
contain.npp <- npp.dat[!is.na(bos.biom30m) & !is.na(lulc), sum(hyb.npp, na.rm=T), by=.(biom.bin, lulc)]
contain.npp <- contain.npp[order(biom.bin), ]
contain.npp[, bin.num:=as.numeric(biom.bin)]
names(contain.npp)[3] <- "tot.npp.kg"
col.tmp <- c("forestgreen", "gray55", "salmon", "gold", "chartreuse3")
plot(contain.npp[lulc==1, bin.num], contain.npp[lulc==1, tot.npp.kg/1000], 
     col=col.tmp[1], pch=15, type="l", lwd=3,
     ylim=c(0, 500), xlab="Biomass density (Mg/ha)", xaxt="n", ylab="Total NPP (Mg-biomass)", main="Hybrid estimates") 
axis(side = 1, at = seq(0,100, by=20), labels = c(0, 10000, 20000, 30000, 40000, 50000))
for(e in 2:5){
  lines(contain.npp[lulc==e, bin.num], contain.npp[lulc==e, tot.npp.kg/1000], col=col.tmp[e], pch=15, type="l", lwd=3)
}
legend(x = 60, y = 400, legend = c("Forest", "Developed", "HD Resid.", "LD Resid.", "Other veg."), fill=col.tmp, bty="n")

## FIA empirical, canopy basis
contain.npp <- npp.dat[!is.na(bos.biom30m) & !is.na(lulc), sum(fia.empir.npp.kg.hw.forest, na.rm=T), by=.(biom.bin, lulc)]
contain.npp <- contain.npp[order(biom.bin), ]
contain.npp[, bin.num:=as.numeric(biom.bin)]
names(contain.npp)[3] <- "tot.npp.kg"
plot(contain.npp[lulc==1, bin.num], contain.npp[lulc==1, tot.npp.kg/1000], 
     col=col.tmp[1], pch=15, type="l", lwd=3,
     ylim=c(0, 500), xlab="Biomass density (Mg/ha)", xaxt="n", ylab="Total NPP (Mg-biomass)", main="FIA (canopy) estimates") 
axis(side = 1, at = seq(0,100, by=20), labels = c(0, 10000, 20000, 30000, 40000, 50000))
for(e in 2:5){
  lines(contain.npp[lulc==e, bin.num], contain.npp[lulc==e, tot.npp.kg/1000], col=col.tmp[e], pch=15, type="l", lwd=3)
}
legend(x = 60, y = 400, legend = c("Forest", "Developed", "HD Resid.", "LD Resid.", "Other veg."), fill=col.tmp, bty="n")

```

