---
title: "Boston 1m analysis"
author: "Drew Trlica"
date: "`r Sys.Date()`"
---
```{r setup, include=F}
library(rmarkdown)
library(knitr)
library(data.table)
library(ggplot2)
library(kableExtra)
opts_knit$set(root.dir = "~/Desktop/FragEVI")
```
Summary of analysis of 12 separate small-area test zones in the Boston city region, 1m data (NDVI in relation to edge position and cover type)


```{r areas_table, echo=F, results='asis'}
b <- read.csv("~/Desktop/FragEVI/processed/bos.zones.summ.csv", stringsAsFactors = F)

### summary table for cover fractions: Zone type area 10 20 30 >30 CanTotal GrassTotal imp nonimp BarrenTotal
e <- round(b[,c(6,8,10,12,13,15,17,19,20)], digits = 2)
p <- b[,2:4]
pe <- cbind(p,e)
pe[,c(8,9,12)] <- pe[,c(8,9,12)]*100
colnames(pe) <- c("Zone", "Type", "Area (ha)", "10m", "20m", "30m", ">30m", "Can %", "Grass %", "imperv.", "nonimperv.", "Barren %")
gor <- kable(pe, "html", caption = "Table 1: Summary of cover in selected test areas, Boston City Limits")
kable_styling(gor, latex_options = "striped", font_size = 12)

```

Table 1 shows total fraction of Canopy, Grass, and "Barren" for each test zone (% scale), along with relative fraction of canopy by edge distance class and impervious vs. non-impervious barren cover (0-1 scale). Test areas included residential areas of varying development density dominated by impervious cover and street trees; two sections of continuous forest (Stonybrook and Allandale reserves); Boston Common (open-grown trees separated by lawn); a golf course and cemetery dominated by open lawn, and an area of coastal wetland and development with few trees (Beachmont). Canopy area in residential areas and parks was dominated by the near-edge fraction (<10m). Other than golf course, cemetery, and coastal Beachmont grass was usually minority fraction of cover. In dense residential zones (Mattapan, Southend, Allston, Roxbury), impervious was dominant fraction of barren cover; in other residential areas (Dorchester, W. Roxbury) non-impervious was a significant minority. Non-forest naturalistic areas (Commons, Cemetery, Beachmont) showed considerable fraction of non-impervious barren cover -- much of which could be water in Commons and Beachmont. There was no significant barren areas in forested areas or golf course.


```{r ndvi_table, echo=F, results='asis'}
### summary tables for NDVI by cover subtype
h <- round(b[,c(5,7,9,11,14,16,18)], digits=4)
bh <- cbind(as.character(b[,2]), h)
colnames(bh) <- c("Zone", "10m.NDVI", "20m.NDVI", "30m.NDVI", ">30m.NDVI", "NDVI.grass", "NDVI.imp", "NDVI.nonimp")
gar <- kable(bh, format = "html", caption = "Table 2: Summary of median NVDI values (1m) in selected test areas, Boston City Limits")
kable_styling(gar,font_size = 12)
```

Table 2 shows NDVI median values for edge and cover classes in different Boston test zones (NDVI=0.000 where no valid pixels available for that class). Edge trees show greater NDVI variability, with generally reduced NDVI compared with more interior trees in the same region. In all areas containing interior edge classes, NDVI increased from canopy edge to interior. Edge (<10m) canopy NDVI was generally lower and more variable in "street" or open-grown trees compared to edge trees in naturalistic areas with greater canopy depth. Grass showed generally lower  NDVI than canopy in the same zone and was more variable across zones than canopy NDVI. Golf course grass NDVI was highest among grass areas examined (excluding forested areas with low grass%), but was higher than edge canopy NDVI in all areas except golf course and forest. Impervious cover tended to show lower NDVI than non-impervious cover, except in areas with known large expanses of water (Stonybrook, Commons).

````{r boxplotting, echo=F, warnings=F, message=F}
## clustered boxplots
## y is data.table imported from extraction steps in EVI.enhance.R
y <- read.csv("~/Desktop/FragEVI/processed/Bos.tests.extract.data.csv")
y <- as.data.table(y)
tests <- as.character(y[,unique(zone.name)])
fill.pal <- c("gray80", "lightgreen", "darksalmon", "orange", "yellow3", "darkgreen", "gray40")
y$zone <- as.factor(y$zone)
y$cov.ed <- as.factor(y$cov.ed)

p <- ggplot(aes(x=zone, y=ndvi, fill=cov.ed), data=y)+
  geom_boxplot(outlier.shape = NA, position = position_dodge(width=.8))+
  scale_fill_manual(values=fill.pal, name="cover class", labels=c("nonimperv", "grass", "10m", "20m", "30m", ">30m", "imperv."))+
  labs(title="Fig. 1: NDVI by edge+cover, 1m", y="NDVI 1m")+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  scale_x_discrete(name = "Test Zone", labels=tests)+
  scale_y_continuous(breaks = seq(from=-0.4, to=0.6, by=0.1), limits = c(-0.4, 0.65))

p
```

Figure 1 (outliers excluded) shows that in zones with a range of edge distance classes NDVI values tend to rise with increasing edge distance, while contracting in variability. Edge (>10m) canopy and grass values were variable between zones, while interior canopy NDVI was more consistent between zones. NDVI tended to be lower and more consistent in impervious cover between areas  (-0.1 to 0.1) compared to non-impervious barren cover, which tended to be somewhat higher (up to 0.2) and more variable between zones.